### 2016-11-06 scater_case_study/timing_macosko.Rmd

---
title: "Timings for using scater with Macosko dataset"
author: Davis McCarthy, Kieran Campbell, Aaron Lun & Quin Wills
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: united
        highlight: tango
        number_sections: false
        code_folding: show
---

```{r setup, include=FALSE}
library("scater")
```

# Introduction

This notebook shows timing results for using scater with the Macosko et al Drop-seq dataset consisting of approximately 45,000 cells. The dataset in its current format was supplied by Vladimir Kiselev.

# Read data

```{r read-data}
counts_mat <- readRDS("../../../../Dropbox/Projects/single-cell-data/macosko2015/macosko_norm.rds")
dim(counts_mat)
```

# Create `SCESet` object

We can make a minimal `SCESet` object from the count matrix.

```{r create-sceset}
cluster <- colnames(counts_mat)
colnames(counts_mat) <- paste0("Cell", 1:ncol(counts_mat))
system.time(
    sce <- newSCESet(countData = counts_mat)
)
```

This object now contains the count data, and the data transformed to log2-counts-per-million.

```{r show-sceset}
rm(counts_mat)
sce$cluster <- cluster
sce
```

# Calculate QC metrics

We do not appear to have ERCC spike-ins in this dataset, so use mitochondrial genes as feature controls.

```{r calc-qc-metrics}
mtgenes <- grepl("MT:", rownames(sce))
system.time(
    sce <- calculateQCMetrics(sce, feature_controls = list(MT = mtgenes))
)
```

We can look at plots of QC metrics:

```{r plot-pheno}
plotPhenoData(sce, aes(x = log10_total_counts, y = total_features, 
                       colour = pct_counts_feature_controls))
```

Such phenotype data plots take little time to produce, even for ~45,000 cells.

# Subsetting

Subsetting a random sample of 10,000 cells from the `SCESet` object only take a matter of seconds, so filtering out unwanted cells will not be a major hurdle.

````{r subsetting}
system.time(
    sce[, sample(ncol(sce), size = 10000)]
)
```

# Automatic outlier detection without output

Automatic outlier detection using PCA on QC metrics does not take too long.

```{r auto-outlier, results='hide'}
system.time(
    sce <- plotPCA(sce, pca_data_input = "pdata", detect_outliers = TRUE,
                   return_SCESet = TRUE)
)
```

# PCA using expression values

PCA using expression values still takes less than a couple of minutes.

```{r pca}
system.time(
    p1 <- plotPCA(sce, colour_by = "outlier")
)
p1
```

We can look at the original clusters defined by the authors in a PCA plot.

```{r pca-cluster}
system.time(
    p2 <- plotPCA(sce, colour_by = "cluster")
)
p2
```

# Automatic outlier detection with output

Automatic outlier detection using PCA on QC metrics does not take too long.

```{r auto-outlier-with-output}
system.time(
    sce <- plotPCA(sce, pca_data_input = "pdata", detect_outliers = TRUE,
                   return_SCESet = TRUE)
)
```